---
title: "hw6"
author: "Matthew Spotnitz"
date: "11/23/2021"
output: github_document
---
```{r}
library(tidyverse)
library(ggplot2)
library(modelr)
library(mgcv)
```
##Problem 1
I will import and view the birthweight dataset.
```{r}
birthweight_df = read.csv("birthweight.csv")
birthweight_df = janitor::clean_names(birthweight_df)
head(birthweight_df)
tail(birthweight_df)
str(birthweight_df)
view(birthweight_df)
```
There are 4342 observations and 20 variables.  Those variables are combinations of integer and numerical formats. I will dnow drop missing values.
```{r}
birthweight_na_df = drop_na(birthweight_df)
str(birthweight_na_df)
```
There were no missing values.  Therefore, it is appropriate to use the original dataframe. I will make a facetwrap that shows a histogram of each variable

```{r}
ggplot(gather(birthweight_df), aes(value)) + geom_histogram(bins = 10) + facet_wrap(~key, scales = 'free_x')
```
After looking at these plots, I intend to convert babysex, frace, mrace, malform, and mrace into factor variables.
```{r}
birthweight_factor_df = birthweight_df %>% mutate (babysex = as.factor(babysex), frace = as.factor(frace), mrace = as.factor(mrace), malform = as.factor(malform), mrace = as.factor(mrace))
str(birthweight_factor_df)
```
I will make scatter plots for numerical variables that may be associated with birthweight.
```{r}
birthweight_factor_df %>% ggplot(aes(x = bhead, y= bwt)) + geom_point()  ##On inspection, there is a corrleation
birthweight_factor_df %>% ggplot(aes(x = blength, y= bwt)) + geom_point() ##On inspection, there is a corrrelation
birthweight_factor_df %>% ggplot(aes(x = delwt, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = gaweeks, y= bwt)) + geom_point() ##There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = menarche, y= bwt)) + geom_point()#There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = fincome, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = mheight, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = momage, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = ppbmi, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = ppwt, y= bwt)) + geom_point()#There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = smoken, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = wtgain, y= bwt)) + geom_point() #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = parity, y= bwt)) + geom_point() #There is not a strong correlation
```
Now I will plot brithweight by non-continuous variables

```{r}
birthweight_factor_df %>% ggplot(aes(x = babysex, y= bwt)) + geom_boxplot()   #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = frace, y= bwt)) + geom_boxplot()   #There is not a strong correlation
birthweight_factor_df %>% ggplot(aes(x = mrace, y= bwt)) + geom_boxplot() #There is not a strong correlation 
birthweight_factor_df %>% ggplot(aes(x = malform, y= bwt)) + geom_boxplot() # There is not a strong correlation
#pnumlbw and pnumsga are exlusively zero values.
```
On the basis of this exploratory data analysis, a baby's head circumference and length at birth have the strongest correlation with birthweight. I will map the residuals and fitted values of these models.
```{r}
fit = lm(bwt ~ bhead + blength, data = birthweight_factor_df)
birthweight_factor_df %>% modelr::add_residuals(fit) %>% ggplot(aes(x=bhead, y = resid)) + geom_point() ##Overall, there is a uniform distribution of residuals
birthweight_factor_df %>% modelr::add_residuals(fit) %>% ggplot(aes(x=blength, y = resid)) + geom_point() ##Overall, there is a uniform distribution of residuals
```
Now I will plot the fitted values 
```{r}
fit = lm(bwt ~ bhead + blength, data = birthweight_factor_df)
birthweight_factor_df %>% modelr::add_predictions(fit) %>% ggplot(aes(x=bhead, y = pred)) + geom_point() ##Overall, there is a direct correlation between baby's head circumference at birth and fitted values
birthweight_factor_df %>% modelr::add_predictions(fit) %>% ggplot(aes(x=blength, y = pred)) + geom_point() ##Overall, there is a direct correlation betwen baby's length at birth and fitted values.
```
Now I will correlate the fitted values with residuals
```{r}
fit = lm(bwt ~ bhead + blength, data = birthweight_factor_df)
birthweight_factor_df  %>% modelr::add_residuals(fit)%>% modelr::add_predictions(fit) %>% ggplot(aes(x=pred, y = resid)) + geom_point() +    labs(
    title = "Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Residuals",
    caption = "Birthweight Data"
  )  + theme_minimal() #Overall, there is not a strong correlation between fitted values and residuals
```
Now I will fit more models. In order to make the model with baby sex run, the input needs to be numeric and not a factor.  Also, I will need to split the data frame into training and testing data.
```{r}
birthweight_df = mutate(birthweight_df, id = row_number())
view(birthweight_df)
train_df = sample_n(birthweight_df, 80)
test_df = anti_join(birthweight_df, train_df, by = 'id')
```
Fit the linear models with the training data
```{r}
fit = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_length_age = lm(bwt ~ blength + gaweeks, data = birthweight_df)
fit_three = lm (bwt ~ bhead + blength + babysex + bhead*blength*babysex, data = birthweight_df)
```
Now I will add model performance
```{r}
train_df %>% 
  gather_predictions(fit, fit_length_age, fit_three) %>% 
  mutate(model = fct_inorder(model))
```
Cross validation
```{r}
cv_df = crossv_mc(birthweight_df, 100)
```
Creating testing and training datasets
```{r}
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```
Now we will do cross validation on different models
```{r}
cv_df = 
  cv_df %>% 
  mutate(
    model_one  = map(train, ~lm(bwt ~ bhead + blength, data = .x)),
    model_two  = map(train, ~lm(bwt ~ blength + gaweeks, data =.x)),
    model_three  = map(train, ~lm (bwt ~ bhead + blength + babysex + bhead*blength*babysex, data = .x))) %>% 
   mutate(
    rmse_model_one = map2_dbl(model_one, test, ~rmse(model = .x, data = .y)),
    rmse_model_two = map2_dbl(model_two, test, ~rmse(model = .x, data = .y)),
    rmse_model_three = map2_dbl(model_three, test, ~rmse(model = .x, data = .y)))
```

Now I will make violin plots of each model
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()+    labs(
    title = "Violin Plot of Residuals vs. Candidate Models",
    x = "Candidate Models",
    y = "Residuals",
    caption = "Birthweight Data"
  )  + theme_minimal()
```
Overall, the model that used child length at birth and gestational age, without interaction terms, had the highest residual error ("model_two).  My model, which used head circumference  and child length at birth without interactions("model_one"), had similar performance to a model that used head cirumferece and length at birth, sex, and interactions terms of all three variables ("model_three").

#Problem Two
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
head_df = janitor::clean_names(weather_df)
head(weather_df)
tail(weather_df)
str(weather_df)
view(weather_df)
```
We are going to fit a linear model with tmax as the response and tmin as the predictor
```{r}
lm (tmax~tmin, data=weather_df) %>% broom::glance()
```
Now I am going to make a boot smaple funcion.
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```
I will run the boot sample function to check that it works
```{r}
boot_sample(weather_df) %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm")
```
I wll now run 5000 boot straps
```{r}
boot_straps = 
  tibble(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weather_df))
  )

boot_straps
```

```{r}
bootstrap_results = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

bootstrap_results %>% 
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975))  
```
On the basis of this output the 95% confidence interval for r squared ranges from 0.894 to 0.927.

```{r}
bootstrap_results = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

bootstrap_results %>%   group_by(term) %>% 
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975))  
```
On the basis of this output the 95% confidence interval of log(B0 • B1) ranges from the log (6.71 * 1.01) to log (7.71 * 1.07). Alternatively, those values are 0.83 to 0.92.