---
title: "hw6"
author: "Matthew Spotnitz"
date: "11/23/2021"
output: github_document
---
```{r}
library(tidyverse)
library(ggplot2)
```
##Problem 1
I will import and view the birthweight dataset.
```{r}
birthweight_df = read.csv("birthweight.csv")
birthweight_df = janitor::clean_names(birthweight_df)
head(birthweight_df)
tail(birthweight_df)
str(birthweight_df)
view(birthweight_df)
```
There are 4342 observations and 20 variables.  Those variables are combinations of integer and numerical formats. I will dnow drop missing values.
```{r}
birthweight_na_df = drop_na(birthweight_df)
str(birthweight_na_df)
```
There were no missing values.  Therefore, it is appropriate to use the original dataframe. I will make a facetwrap that shows a histogram of each variable

```{r}
ggplot(gather(birthweight_df), aes(value)) + geom_histogram(bins = 10) + facet_wrap(~key, scales = 'free_x')
```
After looking at these plots, I intend to convert babysex, frace, mrace, malform, and mrace into factor variables.

